Fractional Types

### 2016-12-18 typespp.tex
removed redundant def
actually there was a redundant section for some reason; fixed
a few thoughts

To understand the challenge in designing an operational semantics for $\eta$ and
$\epsilon$ in the non-dependent setting, we consider the ``zig-zag'' circuit in
Fig.~\ref{fig:zigzag}. Algebraically this circuit corresponds to the following
manipulation on types:
\[
x \mapsto (1 \times x) \mapsto ((x \times 1/x) \times x)
\mapsto (x \times (1/x \times x)) \mapsto (x \times 1) \mapsto x
\]
and by the coherence conditions of compact closed categories, we expect this
circuit, when run in either direction, to be equivalent to the
identity. Consider now the case of a combinator $c$ whose order is greater than
or equal to 2, i.e., the type $\order{c}$ has at least two distinct values
$\idiso$ and $c$ and the type $\iorder{c}$ has at least two distinct values
$1/\idiso$ and $1/c$. We require this circuit to produce $\idiso$ when given
$\idiso$ from either end, and similarly to produce $c$ when given $c$ from
either end. The problem becomes apparent when one observes that the uses of
$\eta+$ in the left-to-right execution and the use of $\epsilon-$ in the
right-to-left execution must behave differently depending on the incoming
value. But the ``local'' information available to either of these combinators
does not include the ``global'' information about the incoming value and there
is no way for them to make consistent guesses locally. It is possible to modify
the operational semantics to use a mechanism to express a computational effect
which enables spatially separated parts of the program to communicate in a way
that is reminiscent of entanglement in quantum mechanics. Possibilities to
realize this communication are global reference cells, backtracking, or logical
variables with unification. Instead, we choose a more abstract approach that
encodes the required dependency in dataflow constraints using dependent types.

\begin{definition}[Dependent $\Pi^/$]
\label{def:pifracd}
The syntax of dependent $\Pi^/$ differs from Def.~\ref{def:pifrac} as follows:
\[\begin{array}{lrcl}
(\textrm{Fractional Types}) & \mathbb{T_\bullet} &::=&
  [\order{c},c^k] \alt
  [\iorder{c},1/c^k] \alt
  [\mathbb{T}_1 \circledast \mathbb{T}_2,\langle v_1,v_2 \rangle] \alt
  \forall k.\mathbb{T_\bullet} \\
 (\textrm{/-combinators}) & \rho_\bullet &:&
  \mathbb{T}_{\bullet 1} \leftrightblackspoon \mathbb{T}_{\bullet 2} ~::=
  [\textit{see Fig.~\ref{pifracd:comb}}]
\end{array}\]
\end{definition}

%%%%%
\subsection{Dependent Pointed Types}

@@

>>>>

@@

Need to explain that a $\forall$ in the forward direction is a choice and
becomes $\exists$ in the backwards direction.

Rules for moving quantifiers?

Combinators for packing (creating an existential) and synchronization
(instantiating a universal)??
